---
title: "Map of Bangladesh from shape file"
author: "Ferdausur Rahman"
date: "12/21/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# __Hi!__  
### I am Ferdausur Rahman


I tried to plot __the map of Bangladesh__ from __.shp__ file. The file can be found on Google by searching with __Bangladesh.shp or Bangladesh shape file__.
I took the file from __[DIVA-GIS](https://www.diva-gis.org/gdata)__ as __.zip__ file and extracted in my working directory. 


## Setting Working Directory
```{r}
wd<- getwd()
```


## Load libraries 
__sf__ package is to read shape file and __ggplot2__ to plot
```{r}
library(sf)
library(ggplot2)

```



## Reading data
The shape file could also be read with __st.read()__ from __sf__ package and __readOGR()__ from __rgdal__ package. 

adm0 contains information on the country, country name.
```{r}
adm0<- read_sf(paste(wd, "BGD_adm/BGD_adm0.shp", sep= "/"))
head(adm0)

```

adm1 contains information on division.
```{r}
adm1<- read_sf(paste(wd, "BGD_adm/BGD_adm1.shp", sep= "/"))
head(adm1)

```

adm2 in on districts of Bangladesh.

```{r}
adm2<- read_sf(paste(wd, "BGD_adm/BGD_adm2.shp", sep= "/"))
head(adm2)

```

adm3 has  previous information and information on upazila.
```{r}
adm3<- read_sf(paste(wd, "BGD_adm/BGD_adm3.shp", sep= "/"))
head(adm3)
```



## Plotting the map  
If the shape data was read by __readOGR()__, map could be also be plotted by getting latitude, longitude from __tidy()__ of __broom__ package and then calling __geom_polygon()__ on the file. 


### A solid map of Bangladesh as adm0 only have country information. 
```{r}
map0<- ggplot(adm0)+
        geom_sf(aes(fill= NAME_0))+
        labs(title = "A solid map of Bangladesh", 
             x= "Longitude", y= "Latitude")+
        theme_classic()+
        theme(legend.title = element_blank())
        
map0

```

### adm1 has divisional distribution on column NAME_1.
```{r}
map1<- ggplot(adm1)+
        geom_sf(aes(fill= NAME_1))+
        labs(title= "A map of Bangladesh with divisional distribution",
             x= "Longitude", y= "latitude")+
        scale_fill_manual("Divisions",
                          values= c("violet", "blue", "cyan", "green", 
                                    "yellow", "orange", "red"))+
        theme_classic()
map1

```

### A map of Bangladesh on districts 
```{r}
map2 <- ggplot(adm2)+
        geom_sf(aes(fill=NAME_1))+
        labs(title="A map of Bangladesh with district distribution",
             x="Longitude",y="Lattitude")+
        theme_classic()+
        scale_fill_manual("Divisions",
                          values= c("violet", "blue", "cyan", "green", 
                                    "yellow", "orange", "red"))
map2


```

### A map of Bangladesh on Upazila
```{r}
map3<- ggplot(adm3)+
        geom_sf(aes(fill= NAME_1))+
        labs(title= "A map of Bangladesh with upazila subdivision", 
             x= "Longitude", y= "Latitude")+
        theme_classic()+
        scale_fill_manual("Divisions",
                          values= c("violet", "blue", "cyan", "green", 
                                    "yellow", "orange", "red"))
map3

```

## Adding labels on the map

Labels can be added both as text and labels.

+ Text and label can be added with __geom_text()__ and __geom_label()__ respectively. However, they need x and y coordinates along with the label/ texts.   
Approximate x and y can be taken from plotting the map without label first and then re-plotting with labels and coordinates.

```{r}
lats<- c(25.800, 24.800, 24.700, 23.600, 23.100, 22.850, 22.500, 21.000)
longs<- c(89.200, 91.600, 89.000, 90.100, 89.400, 91.900, 90.300, 90.800)

label_name<- c("Rangpur", "Sylhet", "Rajshahi", "Dhaka", 
         "Khulna", "Chittagong", "Barishal", "Bay of Bengal")

lab_df<- data.frame(label_name, longs, lats)

```

```{r}
map4 <- ggplot(adm2)+
        geom_sf(aes(fill=NAME_1))+
        labs(title="A map of Bangladesh with divisional distribution",
             x="Longitude",y="Lattitude")+
        theme_classic()+
        geom_label(data = lab_df, 
                   aes(x= longs, y= lats, label= label_name), 
                   alpha= 0.4)+
        scale_fill_manual("Divisions", 
                          values= c("violet", "blue", "cyan", "green", 
                                    "yellow", "orange", "red"))
map4

```

+ Text and labels can be also added with __geom_sf_text()__ and __geom_sf_label()__. They do not require x and y coordinates for adding label/text. 

```{r}
map5 <- ggplot(adm1)+
        geom_sf(aes(fill=NAME_1))+
        labs(title="A map of Bangladesh with divisional distribution",
             x="Longitude",y="Lattitude")+
        theme_classic()+
        geom_sf_label(aes(label= NAME_1), alpha= 0.4)+
        scale_fill_manual("Divisions", 
                          values= c("violet", "blue", "cyan", "green", 
                                    "yellow", "orange", "red"))

map5

```



## Making fill according to a given variable/ column/ information

Population of Bangladesh is taken from [Divisions of Bangladesh](https://en.wikipedia.org/wiki/Divisions_of_Bangladesh) Wikipedia page is added to the data

```{r}
adm1$population<- c(8325666, 29145000, 36433505, 15687759, 
                      18485858, 15787758, 9807000)

```

```{r}
map6 <- ggplot(adm1)+
        geom_sf(aes(fill=population))+
        labs(title="A map of Bangladesh with population distribution",
             subtitle = "by divisions",
             x="Longitude",y="Lattitude")+
        theme_classic()+
        geom_sf_text(aes(label= NAME_1), color= "white")

map6

```


### Thank you for reading. Hope this helps you. 